---
title: "Employee Satisfaction"
author: "Jamie Vo"
date: "4/14/2020"
output: html_document
---

## Abstract
***
Checkout the presentation at: ......youtube video link

## Introduction
***
## Data Description
***
#### Read in the data and load in all the libraries
```{r,warning=FALSE,message=FALSE}
# load in the libraries
library(stringi)
library(plotly)
library(plyr)
library(class)
library(caret)
library(e1071)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(tidyverse)
library(shinydashboard)
library(shiny)
library(glue)
library(reshape)
library(dplyr)
library(FNN) 
library(gmodels) 
library(psych)

# read in the dataframe
data = read.csv("../data_sets/CaseStudy2-data.csv",header = TRUE)
```

## EDA - Exploratory Data Analysis
***
### Clean Data
After checking the data for duplicates and null values, it was determined that the data was clean and no additional 
cleaning would be required.

```{r}
# check for unique employees only
distinct_df<-distinct(data, EmployeeNumber)

# check for complete observations 
df[complete.cases(data), ]

```

### Explore Data

#### Set the datasets 
##### categorical variables
Attrition, BusinessTravel, Department, EducationField, Gender, JobRole, MaritalStatus, OverTime, JobLevel, JobSatisfaction, JobInvolvement, PerformanceRating, RelationshipSatisfaction,StandardHours, StockOptionLevel, NumCompaniesWorked, TrainingTimesLastYear, WorkLifeBalance, Education, YearsInCurrentRole, EnvironmentSatisfaction

##### continuous variables
ID, Age, Attrition, DailyRate, DistanceFromHome, EmployeeNumber, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, TotalWorkingYears, YearsAtCompany, YearsSinceLastPromotion, YearsWithCurrManager
```{r}
# create a dataframe with categorical, removing Over 18 and 
categorical <- subset(data, select = c(Attrition, BusinessTravel, Department, EducationField, Gender, JobRole, MaritalStatus, OverTime, JobLevel, JobSatisfaction, JobInvolvement, PerformanceRating, RelationshipSatisfaction, StockOptionLevel, NumCompaniesWorked, TrainingTimesLastYear, WorkLifeBalance, Education, YearsInCurrentRole, EnvironmentSatisfaction))

# create a dataframe with all continuous
continuous <- subset(data, select = c(ID, Age, Attrition, DailyRate, DistanceFromHome, EmployeeNumber, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, TotalWorkingYears, YearsAtCompany, YearsSinceLastPromotion, YearsWithCurrManager, StandardHours) )

```


#### Determine relationships to Attrition

##### Numerical relations
Reviewing the plots below, the numerical factors that appear to correlate with higher attrition rates is:

* Monthly Income
* Percent Salary Hike
* Total Working Years
* Years at the Company
* Years Since Last Promotion

The were selected due to the clustering or correlation R2 of positive attrition.
```{r,warning=FALSE,message=FALSE}
# differentiate colors by attrition
my_cols <- c("#00AFBB", "#E7B800") 

df_attrition_no <- continuous %>% filter(Attrition == "No")
df_attrition_yes <- continuous %>% filter(Attrition == "Yes")

# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 1, col = my_cols[continuous$Attrition])
}
# Create the plots
#pairs(continuous[,1:5], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business
#pairs(continuous[,5:14], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business
#pairs(continuous[,10:14], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business

# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 1, col = my_cols[df_attrition_yes$Attrition])
}
pairs(df_attrition_yes[,1:5], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business
pairs(df_attrition_yes[,5:10], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business
pairs(df_attrition_yes[,10:15], lower.panel = panel.cor, upper.panel = upper.panel) # age vs. business
```

##### Categorical Relations
It's apparent that only specific variables are significant in terms of attrition. They variables are listed below:
* *Job Role
* *Marital Status
* *Overtime
* Job Level
* *Job Involvement
* Stock Option Level
These variables has the most stark differences when it comes to the percentage differences between the variables.

It is interesting to note that gender, education, and job satisfaction didn't have a major impact on attrition.

```{r}

#compare the plots with attrition being the explainatory factor (Categorical)
data$AttritionF <- as.factor(data$Attrition)
categorical_variables = colnames(categorical)

graphing <- function(data, categorical_variables) {
  for (val in categorical_variables) {
    plot <- ggplot(data=data)+aes_string(val)+geom_bar(aes(fill=as.factor(AttritionF)), position="fill")+ ggtitle(glue("{val}"))
    print(plot)
  }
}
graphing(data, categorical_variables)



``` 
##### Test variables for relevance
Pull the variables visually analyzed and test for relevance in the final model. Since we are testing for a categorical variable, KNN, is utlized. 


##### ETL 
Scaling the data to prevent loss of patterns that would not been seen if not completed. Extract the data that's determined the most relevant from the analysis above.
```{r}
# Scale the data

# scale the data
data$MonthlyIncomeScaled <- scale(data$MonthlyIncome)
data$PercentSalaryHikeScaled <- scale(data$PercentSalaryHike)
data$TotalWorkingYearsScaled <- scale(data$TotalWorkingYears)
data$YearsAtCompanyScaled <- scale(data$YearsAtCompany)
data$YearsSinceLastPromotionScaled <- scale(data$YearsSinceLastPromotion)

# grab all significant variables for testing
variables_graphed <- subset(data, select = c(ID, Attrition, JobRole, MaritalStatus, OverTime, JobLevel, JobInvolvement, StockOptionLevel, WorkLifeBalance, MonthlyIncome, PercentSalaryHike, TotalWorkingYears, YearsAtCompany, YearsSinceLastPromotion, MonthlyIncomeScaled, PercentSalaryHikeScaled, TotalWorkingYearsScaled, YearsAtCompanyScaled, YearsSinceLastPromotionScaled))

need_dummycode <- c("JobRole", "MaritalStatus", "OverTime")
for (val in need_dummycode) {
  temp <- as.data.frame(dummy.code(data[[val]]))
  variables_graphed <- cbind(variables_graphed, temp)
}
colnames(variables_graphed)
```


Its determined that K-15 results is the best KNN-model. 
* Monthly Salary
* Overtime
* Role

at K = 7, seed = 7
```{r}
# cross validation knn
set.seed(7)
splitPerc = .80

#split data
trainIndices = sample(1:dim(variables_graphed)[1],round(splitPerc * dim(variables_graphed)[1]))
train = variables_graphed[trainIndices,]
test = variables_graphed[-trainIndices,]

# create dataframe to hold KNN results
accs = data.frame(accuracy = numeric(20), k = numeric(20))

# run KNN algorithm 
for(i in 1:20)
{
  classifications = knn(train[,c(32,33,15,20,21,22,23,24,25,26,27,28)],test[,c(32,33,15,20,21,22,23,24,25,26,27,28)],train$Attrition, prob = TRUE, k = i)
  #print(table(test$Attrition,classifications))
  tryCatch({ CM <- confusionMatrix(table(test$Attrition,classifications))
    accs$accuracy[i] = CM$overall[1]
    accs$specificity[i] = CM$byClass[2]
    accs$sensitivity[i] = CM$byClass[1]
    accs$k[i] = i}, 
    error = function(e) {
    accs$accuracy[i] = 0
    accs$specificity[i] = 0
    accs$sensitivity[i] = 0
    accs$k[i] = 0
    })
  #CM <- confusionMatrix(table(test$Attrition,classifications))
}
accs
plot(accs$k,accs$accuracy, type = "l", xlab = "k", main="K Iterations and Accuracy") 

colnames(variables_graphed)

train[,c(1,2,3)]
train[,c(1,2,7)]
```

##### Test KNN against different seeds

```{r}

accs1 = data.frame(accuracy = numeric(20), k = numeric(20))
# run KNN algorithm 
for(i in 1:20)
{
  # set seed
  set.seed(i)
  splitPerc = .80
  
  
  #split data
  trainIndices = sample(1:dim(variables_graphed)[1],round(splitPerc * dim(variables_graphed)[1]))
  train = variables_graphed[trainIndices,]
  test = variables_graphed[-trainIndices,]
# run KNN algorithm 

  classifications = knn(train[,c(32,33,15,20,21,22,23,24,25,26,27,28)],test[,c(32,33,15,20,21,22,23,24,25,26,27,28)],train$Attrition, prob = TRUE, k = 7)
  #print(table(test$Attrition,classifications))
  tryCatch({ CM <- confusionMatrix(table(test$Attrition,classifications))
    accs1$accuracy[i] = CM$overall[1]
    accs1$specificity[i] = CM$byClass[2]
    accs1$sensitivity[i] = CM$byClass[1]
    accs1$seed[i] = i}, 
    error = function(e) {
    accs1$accuracy[i] = 0
    accs1$specificity[i] = 0
    accs1$sensitivity[i] = 0
    accs1$seed[i] = 0
    })
  #CM <- confusionMatrix(table(test$Attrition,classifications))
}
accs1
summary(accs1)

```


## Data Conclusions
***
### Attrition
Attrition is defined as reducing/preventing voluntary employee turnover. Here, if the value is no, this means that the employee was not lost, and yes means that the company did lose the employee. 

#### Factors that Lead to Attrition
While many factors were considered to lead to attrition, it was determined that the top 3 factors are 
1. Monthly Income
2. Over Time
3. Job Role

#### Model for Attrition Prediction
The final model takes into account all three factors, with a mean specificity of 0.61 and sensitivity of 0.87 at K-7.
```{r}
# select desired columns from the data sets 
training_set <- train %>% select(c("OverTime", "JobRole", "MonthlyIncome"))
testing_set <- test %>% select(c("OverTime", "JobRole", "MonthlyIncome"))

#since the data needs to be dummy coded for categorical, the following is more accurate
training <- train[,c(32,33,15,20,21,22,23,24,25,26,27,28)]
testing<- test[,c(32,33,15,20,21,22,23,24,25,26,27,28)]

# run the data
CM <- knn(training,testing,train$Attrition, prob = TRUE, k = 7)

# save the column to the desired dataset
testing$prediction<- CM


```

##### Predicting Attrition
At minimum, the attrition should be at 60% and the sensitivity should be 60%.
Aftering running the model to recieve predicted attrition values, the dataset was ordered by id and exported to a csv. 
```{r}
# load in the attrition data
attrition_competition = read.csv("../data_sets/CaseStudy2CompSet No Attrition.csv", header=TRUE)


# order by id
attach(attrition_competition)
attrition_competition_ordered <- attrition_competition[order(ID),]
colnames(attrition_competition_ordered)
# run predictions on the competition model

# 1. dummy code data
need_dummycodes <- c("JobRole", "OverTime")
for (val in need_dummycodes) {
  temp <- as.data.frame(dummy.code(attrition_competition_ordered[[val]]))
  attrition_competition_ordered <- cbind(attrition_competition_ordered, temp)
}

# 2. Scale the monthly income
attrition_competition_ordered$MonthlyIncomeScaled <- scale(attrition_competition_ordered$MonthlyIncome)

# 3. pull the train and validation sets
training <- train[,c(32,33,15,20,21,22,23,24,25,26,27,28)]
testing <- attrition_competition_ordered %>% select(c("Research Scientist","Sales Executive","Laboratory Technician","Manufacturing Director","Manager","Healthcare Representative","Research Director","Sales Representative","Human Resources","No","Yes","MonthlyIncomeScaled"))

# 4. Run the model to get predictions
CM <- knn(training,testing,train$Attrition, prob = TRUE, k = 7)

# 5. save the column to the desired dataset
attrition_competition_ordered$Attrition<- CM

# 6. filter for desired columns
output_attrition <- attrition_competition_ordered %>% select(c("ID", "Attrition"))
# export the attrition predicted data
write.csv(output_attrition, "../data_sets/output_data_sets/Case2Predictions_JamieVo Attrition.csv")

```

### Monthly Incomes

#### Model for Monthly Incomes
```{r}

```


##### Predicting Monthly Incomes
The model attains a RMSE < $3000 for the training and the validation set.
```{r}
# load in the monthly income data
#monthly_income_competition = read.csv("../data_sets/CaseStudy2CompSet No Salary.csv", header=TRUE)


# order by id
#attach(attrition_competition)
#monthly_income_competition_ordered <- monthly_income_competition[order(ID),]

# export the attrition predicted data
#write.csv(monthly_income_competition_ordered, "../data_sets/output_data_sets/Case2Predictions_JamieVo Salary.csv")
```


#### Job Specific Trends
```{r}

```

#### Additional Trends
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
